---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, 
               cache = FALSE,
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Tango

Chaccour and Brew


<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
<!-- badges: end -->


```{r}
# Load the package for this project
library(tango)
# Load other useful packages
library(knitr)
library(tidyr)
library(dplyr)
library(sp)
library(ggplot2)
library(RColorBrewer)

# Get census age data
census <- tango::census
# Get municipios spatial data
municipios <- tango::municipios


# Get census age data
census <- diadespres::census
#43 = Tarragona
#08 = Barcelona
#25 = Lleida
#17 = Girona
catalan_codes <- c('43', '08', '25', '17')
census <- census %>%
  filter(substr(id, 1, 2) %in% catalan_codes)

# Get municipios spatial data
municipios <- diadespres::municipios
municipios <- municipios[substr(municipios@data$id, 1, 2) %in% catalan_codes, ]

# Order census by pop
by_pop <- census %>%
  group_by(id, municipio) %>%
  summarise(pop = sum(total, na.rm = T)) %>%
  ungroup %>%
  arrange(desc(pop)) %>%
  mutate(p = pop / sum(pop) * 100) %>%
  mutate(cp = cumsum(p))

areas <- rgeos::gArea(municipios, byid = T)

# Make plot by percentage
make_plot <- function(cut_off = 30){
  sub_pop <- by_pop %>%
    filter(cp <= cut_off)
  keep <- municipios@data$id %in% sub_pop$id
  keep_areas <- areas[keep]
  sub_municipios <- municipios[keep,]
  sub_municipios@data$area <- keep_areas
  space_take_up <- round(sum(keep_areas) / sum(areas) * 100, digits = 1)
  plot(municipios)
  plot(sub_municipios, col = 'red', add = T)
  title(main = paste0(max(round(sub_pop$cp)), '% of Catalans live in the red area (', space_take_up, ' % of area)' ))
}

# for(i in c( seq(30, 90, by = 10), 92, 94, 96, 98, 99)){
#   make_plot(cut_off = i)
# }

```

## Plots for 2020-04-06

### Population density

```{r}
# # Project municipalities
# municipios_proj <- spTransform(municipios, CRS("+proj=lcc +lat_1=40 +lat_0=40 +lon_0=0 +k_0=0.9988085293 +x_0=600000 +y_0=600000 +a=6378298.3 +b=6356657.142669561 +pm=madrid +units=m +no_defs"))

# Get area for each municipality
areas_km <- raster::area(municipios) / 1000000

# Add to municipios
map <- municipios
map@data$area_km <- areas_km

# Get population numbers to plug into municipios
right <- census %>%
  group_by(id) %>%
  summarise(pop = sum(total, na.rm = TRUE),
            pop60 = sum(total[edad >= 60], na.rm = TRUE),
            pop50 = sum(total[edad >= 50], na.rm = TRUE),
            pop70 = sum(total[edad >= 70], na.rm = TRUE),
            pop80 = sum(total[edad >= 80], na.rm = TRUE),
            pop90 = sum(total[edad >= 90], na.rm = TRUE),
            popl60 = sum(total[edad < 60], na.rm = TRUE),
            popl50 = sum(total[edad < 50], na.rm = TRUE),
            popl70 = sum(total[edad < 70], na.rm = TRUE),
            popl80 = sum(total[edad < 80], na.rm = TRUE),
            popl90 = sum(total[edad < 90], na.rm = TRUE))
  
map@data <-
  left_join(map@data,
            right)

map@data <- map@data %>%
  mutate(pop_per_km = pop / area_km,
         pop50_per_km = pop50 / area_km,
         pop60_per_km = pop60 / area_km,
         pop70_per_km = pop70 / area_km,
         pop80_per_km = pop80 / area_km,
         pop90_per_km = pop90 / area_km,
         popl50_per_km = popl50 / area_km,
         popl60_per_km = popl60 / area_km,
         popl70_per_km = popl70 / area_km,
         popl80_per_km = popl80 / area_km,
         popl90_per_km = popl90 / area_km) %>%
  mutate(ratio60 = popl60 / pop60,
         ratio50 = popl50 / pop50,
         ratio70 = popl70 / pop70,
         ratio80 = popl80 / pop80,
         ratio90 = popl90 / pop90)
big_right <- map@data

# Fortify
mapf <- fortify(map, region = 'id')

library(ggthemes)
```

# 50 -/+ ratio

```{r}
map@data$ratiox <- ifelse(map@data$ratio50 < 1,
                          '0-1',
                          ifelse(map@data$ratio50 <2,
                                 '1-2',
                                 ifelse(map@data$ratio50 < 3,
                                        '2-3',
                                        ifelse(map@data$ratio50 < 4,
                                               '3-4',
                                               '4+'))))

shp <- mapf %>%
  left_join(map@data[,c('id', 'ratiox')])

cols <-  rev(RColorBrewer::brewer.pal(n = 5, name = 'Spectral'))
ggplot(data = shp,
         aes(x = long,
             y = lat,
             group = group,
             fill = factor(ratiox))) +
    geom_polygon(color = 'black',
         size = 0.2) +
         theme_map() +
    theme(plot.title = element_text(size = 16),
          legend.position = 'right') +
        scale_fill_manual(name ='',
                         values  = cols) +
  labs(title = 'Ratio of people younger than 50 to people older than 50')
```


Table of cut-offs and population affected. How to read:

-Cut off: We apply measures to all towns with this ratio, or lower, of young to old
-Below: Means that the measures are applied
-Above: means that the measures are not applied.

```{r}
pop <- sum(map@data$pop)
out_list <- list()
cut_offs <- c(1, 1.5, 1.7, 1.8, 1.9, 2, 3, 4, 5)
for(i in 1:length(cut_offs)){
  this_cut_off <- cut_offs[i]
  out <- data.frame(cut_off = this_cut_off)
  out$under_50s_affected <- sum(map@data$popl50[map@data$ratio50 < this_cut_off])
  out$over_50s_protected <- sum(map@data$pop50[map@data$ratio50 < this_cut_off])
  out$pop_below_cut_off <- sum(map@data$pop[map@data$ratio50 < this_cut_off])
  out$pop_above_cut_off <- sum(map@data$pop[map@data$ratio50 >= this_cut_off])
  out$p_below_cut_off <- round(out$pop_below_cut_off / pop * 100, digits = 2)
  out$p_above_cut_off <- round(out$pop_above_cut_off / pop * 100, digits = 2)
  out$municipalities_below_cut_off <- length(map@data$pop[map@data$ratio50 < this_cut_off])
  out$municipalities_above_cut_off <- length(map@data$pop[map@data$ratio50 >= this_cut_off])
  out_list[[i]] <- out
}
out <- bind_rows(out_list)
knitr::kable(out)
```


```{r}
cut_pretty <- function(x, breaks, collapse=" to ", ...) {
  breaks[1] <- floor(breaks[1])
  breaks[length(breaks)] <- ceiling(breaks[length(breaks)]) 
  breaks[2:(length(breaks)-1)] <- round(breaks[2:(length(breaks)-1)])
  breaks <- unique(breaks)
  it_breaks <- itertools2::ipairwise(breaks)
  breaks_pretty <- sapply(it_breaks, paste, collapse=collapse)
  cut(x, breaks=breaks, labels=breaks_pretty, ...)
}

# Define variable for plotting var
plot_var <- function(var = 'pop', return_table = 0, quant = T, n = 5){
  sub_map <- mapf
  right <- big_right[,c('id', var)]
  sub_map <- left_join(sub_map, right)
  names(sub_map)[ncol(sub_map)] <- 'var'
  if(return_table > 0){
    out <- big_right[,c('id', 'NAMEUNIT', var)]
    names(out)[ncol(out)] <- 'var'
    out <- out %>% arrange(desc(var))
    names(out)[ncol(out)] <- var
    out$pop <- big_right$pop
    names(out)[2] <- 'Nom'
    out <- out[1:return_table,]
    out
  } else {
    if(quant){
      x <-cut_pretty(sub_map$var, breaks = unique(unique(quantile(right[,2], probs = seq(0, 1, length = n)))))
      sub_map$var <- x
      cols <-  rev(RColorBrewer::brewer.pal(n = 4, name = 'Spectral'))
       ggplot(data = sub_map,
         aes(x = long,
             y = lat,
             group = group,
             fill = factor(var))) +
    geom_polygon(color = 'black',
         size = 0.2) +
         theme_map() +
    theme(plot.title = element_text(size = 16),
          legend.position = 'right') +
        scale_fill_manual(name ='',
                         values  = cols)
    } else {
      cols <-  rev(RColorBrewer::brewer.pal(n = 9, name = 'Spectral'))
      ggplot(data = sub_map,
         aes(x = long,
             y = lat,
             group = group,
             fill = var)) +
    geom_polygon(color = 'black',
         size = 0.2) +
            theme_map() +
    theme(plot.title = element_text(size = 16),
          legend.position = 'right') +
        scale_fill_gradientn(name ='',
                         colors  = cols)
    }
  }
}
```

# Population

```{r}
plot_var('pop') +
  labs(title = 'Population per municipality')
```


# Population density

```{r}
plot_var('pop_per_km') +
  labs(title = 'Population per sq km')
```

Top 10:

```{r}
plot_var('pop_per_km', return_table = 10)
```


# Population density of 50+ year-olds

```{r}
plot_var('pop50_per_km') +
  labs(title = 'Population of people aged 50+ per sq km')
```

Top 10:

```{r}
plot_var('pop50_per_km', return_table = 10)
```



# Population density of 60+ year-olds

```{r}
plot_var('pop60_per_km') +
  labs(title = 'Population of people aged 60+ per sq km')
```

Top 10:

```{r}
plot_var('pop60_per_km', return_table = 10)
```


# Population density of 70+ year-olds

```{r}
plot_var('pop70_per_km') +
  labs(title = 'Population of people aged 70+ per sq km')
```

Top 10:

```{r}
plot_var('pop70_per_km', return_table = 10)
```


# Population density of 80+ year-olds

```{r}
plot_var('pop80_per_km') +
  labs(title = 'Population of people aged 80+ per sq km')
```

Top 10:

```{r}
plot_var('pop80_per_km', return_table = 10)
```



# Ratio of under-50s to 50+

(ie, ratio of protectors to protected)

```{r}
plot_var('ratio50', n = 10) +
  labs(title = 'Ratio of under-50s to 50+')
```

Top 10:

```{r}
plot_var('ratio50', return_table = 10)
```



# Ratio of under-60s to 60+

(ie, ratio of protectors to protected)

```{r}
plot_var('ratio60') +
  labs(title = 'Ratio of under-60s to 60+')
```

Top 10:

```{r}
plot_var('ratio60', return_table = 10)
```



# Ratio of under-70s to 70+

(ie, ratio of protectors to protected)

```{r}
plot_var('ratio70') +
  labs(title = 'Ratio of under-70s to 70+')
```

Top 10:

```{r}
plot_var('ratio70', return_table = 10)
```


# Ratio of under-80s to 80+

(ie, ratio of protectors to protected)

```{r}
plot_var('ratio80') +
  labs(title = 'Ratio of under-80s to 80+')
```

Top 10:

```{r}
plot_var('ratio80', return_table = 10)
```
